---
title: "The Numbers behind Popular Music"
output: 
  github_document:
    number_section: FALSE
---

# The Numbers behind Popular Music {.unnumbered}

Authors: Jasper Jansen, Jolie Jacobus, Soren Wagstrom, Quita Jackson

## Abstract (TL;DR) {.unnumbered}

Our team investigated the factors behind popular music. We used a dataset detailing the top 5000 albums of all time. This data comes from the website rateyourmusic.com, however we found this dataset on Kaggle.com. Our team was interested in this topic as music fans, and we wanted to explore the trends in popular music over the years. Our main findings were that most higher rated albums were released before 1980. The most rated and reviewed artists on the list were Radiohead, The Velvet Underground, Nirvana, and Pink Floyd. There was no significant trend in ratings of genres over time. Of the most common genres, the highest rated genre was Film Score, and the lowest rated genre was Hard Rock. When the album was released had a negative correlation with the rating for the album, and the number of review and ratings had a positive correlation with the rating of the album.

# Intro/Background/Motivation

Our team investigated the factors behind popular music. We used a dataset detailing the top 5000 albums of all time. This data comes from the website rateyourmusic.com, however we found this dataset on Kaggle.com. This project is relevant because by analyzing trends in popular music, this project could help artists predict future trends in the music industry or help them to decide when to release their music to get the best results.

In section [Results 1] we show the relationship between the album release date and the rating of the album. In section [Results 2], popularity of the artist is considered with the ratings of the albums. In section [Results 3], trends in the ratings of genres is explored over time. In section [Results 4], the relationship between score and the album's genre, descriptors, and artist is investigated.

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = TRUE)


knit_hooks$set( fig.cap = function(before, options, envir) {
  if ( before ) {
    chr <- sprintf("<p><small><strong><a name='fig:%s' value = %d>Figure %d</a></strong>: %s</small></p>", options$label, options$fig.num, options$fig.num, options$fig.cap)
    chr <- sprintf("<p><small><strong><a name='fig:%s'>%s</a></strong>: %s</small></p>", options$label, options$label,  options$fig.cap)
    
    return(chr)
  }
})


chunkref <- local({
  function(chunklabel) {
    sprintf('[%s](#%s)', chunklabel, paste0("fig:",chunklabel) )
  }  
})

library(ggplot2)
library(dplyr)
library(tidyverse)
library(stringr)
library(lubridate)
```

# Quick Data Summary

Our dataset is the list of the top 5000 rated albums on the website rateyourmusic.com. The dataset includes the ranking, album title, artist, release date, genre(s), descriptors, average rating, number of ratings, number of reviews

+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Variable                | Description                                                                                                                                      | Summary Statistics      |
+=========================+==================================================================================================================================================+=========================+
| Ranking                 | The ranking out of 5000 for the album                                                                                                            | Integer type            |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 1-5000           |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Album                   | The Album Name                                                                                                                                   | Character type          |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Artist.Name             | The artist of the album's name                                                                                                                   | Character type          |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Average.Rating          | The average rating of the album                                                                                                                  | Numerical type          |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 3.52-4.34        |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Number.of.Ratings       | The number of ratings the album has                                                                                                              | Integer type            |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 260-70382        |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Number.of.Reviews       | the number of reviews the album has. A review includes a text review, whereas with a rating a user only inputs a numerical rating for the album. | Integer type            |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 0-1549           |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Genre_1 - Genre_5       | The genre the album is classified as. Each album could have up to 5 genre classifications.                                                       | Character type          |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Desc_1 - Desc_10        | The descriptor the album is classified as. Each album could have up to 10 descriptors.                                                           | Character type          |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Artist.Average.Score    | The average rating for all of an artist's albums on the list                                                                                     | Numerical type          |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 3.52-4.34        |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Artist.Average.Reviews  | The average number of reviews an artist's album has.                                                                                             | Numerical type          |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 0-929            |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Artist.Average.Ratings  | The average number of ratings an artist's album has.                                                                                             | Numerical type          |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 260-46438        |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Artist.Frequency        | How many times an artist's album appears in the dataset.                                                                                         | Integer Type            |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Range: 1-25             |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+
| Year, Month, Day        | The year, month and day the album was released of the album.                                                                                     | Integer/Character Type  |
|                         |                                                                                                                                                  |                         |
|                         |                                                                                                                                                  | Year Range: 1947 - 2021 |
+-------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+

```{r}
df_final <- read.csv("df_final.csv")

nrow(unique(as.data.frame(df_final$Album)))
  #Some albums have the same names
nrow(unique(as.data.frame(df_final$Artist.Name)))
  #2787 unique artists
nrow(unique(as.data.frame(df_final$Genre_1)))
  #453 unique primary genres
summary(df_final)
str(df_final)
```

# Results

In your write-up, make sure to refer to all of the figures you create. You can include a hyperlink to the [scatterplot](#fig:scatterplot) by using the name of the code chunk (make sure, to give each code chunk a different name). In your markdown document you can create this link either by calling the function `chunkref` with the name of the code chunk in quotes, i.e. `r chunkref("scatterplot")` or by using the markdown expression `[scatterplot](#fig:scatterplot)`. Similarly, we can refer to the `r chunkref("2nd scatterplot")`. Note that the figure captions appear above the figures - this saves us from having to scroll up after following the link.

```{r scatterplot, echo=FALSE, fig.cap="This is the figure caption. Make sure to use the description we practised in the homework: first sentence describes structure of the plot, second sentence describes main finding, third sentence describes outliers/follow-up.", message = FALSE}

```

```{r 2nd scatterplot, echo=FALSE, fig.cap="This is the figure caption. Make sure to use the description we practised in the homework: first sentence describes structure of the plot, second sentence describes main finding, third sentence describes outliers/follow-up.", message = FALSE}

```

Additionally, you can also refer to different sections in your writeup by using anchors (links) to section headers. Here, we are referring to subsection [Results 3]. The code for that is `[Results 3]`.

## Results 1 {data-link="Results 1"}

-What is the relationship between the release date and the score of the album? **Quita**

```{r}
ggplot(df_final, aes(x=`Year` , y =`Average.Rating`, color= `Year` ))  + geom_point(size=1.25) + theme(axis.text.x = element_text(angle = 90, size = 10)) +theme(legend.position="none")
```
```
This graph shows the average score of each album by the year it was realeased and then colered it by the year so I to better visuaize it.  By doing this, I hoped to figure out which years has the best or worst scored albums. Most of the scores sit between 3.6 and 4. When looking at the graph, I see that the lowest rated album was released in 2017 whereas the best rated album was realeased in 1963. It is important to remember that these are the top rated albums on the site overall, so the ratings are not terrible but it looks like as the years increase, you see the cluster of rating moving more towards the lower end of the septrum. 

```{r}
df_final%>%group_by(Year)%>% summarise(
  Average.Score =mean(Average.Rating)
)%>% ggplot(aes(x=Year,y=Average.Score))+geom_point()+ theme(axis.text.x = element_text(angle = 90, size = 10))
```

This graph scores the average score that all albums recieved by the the year. My conclusion of a overall decresing score as the move from year to year is confirmed when looking at this graph.There is a huge increase in average score as you move towards the 70's but then it starts to decline shortly after. There is an outlier in 1954 but I am choosing to ignore that because the amount of albums was very limited for that year, there is a total of 6 albums released that year in the data set. 


```{r}
df_final%>% filter(
  !(Month=="")) %>%group_by(Month)%>%summarise(
  Average.Score = mean(Average.Rating),
  Year = Year
  )%>%  mutate(Month = factor(Month, levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep", "Oct","Nov","Dec")))%>%ggplot(aes(x=Month,y=Average.Score))+geom_point()
```
This graph looks at the average score of albums by the month they were released it. Music is sometimes seasonal so uing this graph, I might be able to spot potential seasonal trends that may have occured with the Top 5000 Albums. When looking at the graph, it shows that December is the month with the highest rated albums where as May is the month with the lowest rated albums. I would have assumed the complete oppisite since May is right before the summer and that is often the start of festival season but this graph tells me somethind different.



```{r}
df_final%>% filter(
  !(Month==""),!(Day==""))%>% group_by(Day)%>%summarise(
  Average.Score = mean(Average.Rating)
  )%>%ggplot(aes(x=Day,y=Average.Score))+geom_point()
```

This graph shows the realtionhip between the day of realease and the album score. There is not much pattern to this graph but it appears that albums realeased during the month are more highly rated. 


Comclusion: 
Overall, I would say that there is a pattern of when the album is realeased to the score it gets. I cannot make any further conclusions since I do not know more about the audience that is rating the album and did not account for other facors like genre and artist. 




## Results 2 {data-link="Results 2"}

-What is the relationship between the number of reviews/ratings and the artist? This helps determine the relative popularity of artists on this website.

**Soren** \# Question 2: Who are the most well-known artists on this website, and does that correlate with high scores?

Answering this question will involve determining the average number of ratings/reviews per artist. This can be accomplished using mutate statements to create two new columns, one for average number of ratings per artist, and one for average number of reviews per artist. The number of reviews per artist will be much lower than the number of ratings, because writing a review takes much more time and effort than simply providing a score from 1-5. This is not a reflection of an artist's popularity outside the context of this website, as this is a website for dedicated fans of music, who tend to have different tastes than the general public.

The average number of ratings and reviews per artist from each of their appearances on this list is already determined by the data processing above, so the only step remaining for this question is to visualize and communicate the results.

The plot below shows the top 10 artists by average number of ratings. Radiohead tops the list, followed by The Velvet Underground and Nirvana. These are all very popular mainstream artists, so there is no surprise here. They all have very high average scores as well, indicating that these artists are as beloved as they are famous. There are some less well-known artists on this list as well, such as Slint and Godspeed You, Black Emperor!, which may indicate that these artists are more popular among music enthusiast crowds like the RateYourMusic community than among the general public.

```{r}
# Artist popularity by ratings
df_sorted_ratings = df_final[order(-df_final$Artist.Average.Ratings),]
df_sorted_ratings = df_sorted_ratings[,c(3, 25, 27)]
df_sorted_ratings = unique(df_sorted_ratings)[1:10,]
df_sorted_ratings %>%
  ggplot(aes(x = reorder(Artist.Name, Artist.Average.Ratings), 
             y = Artist.Average.Ratings, fill = Artist.Name)) + 
  geom_bar(stat="identity") + coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, size = 10, hjust = 1)) +
  theme(legend.position="none") + 
  labs(x = "Artist Name", y = "Number of Ratings") + 
  geom_text(aes(label=round(Artist.Average.Score, digit = 2), hjust = 1)) 
```

The next plot, shown below, includes the top 10 artists by average number of reviews. Again, The Velvet Underground and Nirvana are in the top 3, but Pink Floyd takes the third spot this time around, followed closely by Radiohead.

There are two outliers in the top 10 in terms of score here. Both Guns N' Roses and The Sex Pistols have a much lower average score than the others on this list. One conclusion that could be drawn from this statistic it that while these two artists are clearly very well known, they are not as widely beloved among music enthusiasts as others on this list, and their fame may somewhat overrate the quality of their music.

```{r}
# Artist popularity by reviews
df_sorted_reviews = df_final[order(-df_final$Artist.Average.Reviews),]
df_sorted_reviews = df_sorted_reviews[,c(3, 25, 26)]
df_sorted_reviews = unique(df_sorted_reviews)[1:10,]
df_sorted_reviews %>%
  ggplot(aes(x = reorder(Artist.Name, Artist.Average.Reviews), 
             y = Artist.Average.Reviews, fill = Artist.Name)) + 
  geom_bar(stat="identity") + coord_flip() + 
  theme(axis.text.x = element_text(angle = 90, size = 10, hjust = 1)) +
  theme(legend.position="none") + 
  labs(x = "Artist Name", y = "Number of Reviews") + 
  geom_text(aes(label=round(Artist.Average.Score, digit = 2), hjust = 1))
```


## Results 3 {data-link="Results 3"}

# Question: How do scores for genres change over time?

```{r}
length(unique(df_final$Genre_1))
```

There are 435 unique genres in the dataset. This would be too much to cover in a plot, so we will focus on the genres that occur most on the list.

```{r}
df_final %>% group_by(Genre_1)%>% summarise(
  Average.Score = mean(Average.Rating),
  count = n()
)%>% arrange(desc(count))
```

Progressive Rock, Singer/Songwriter, Hard Bop, Film Score, Hard Rock, Heavy Metal, and East Coast Hip Hop are the top seven genres the occur the most on the list.

```{r}
df_final%>% filter(Genre_1 == "Progressive Rock"|
            Genre_1 == "Singer/Songwriter"|
            Genre_1 == "Hard Bop"|
            Genre_1 == "Film Score"|
            Genre_1 == "Hard Rock"|
            Genre_1 == "Heavy Metal"|
            Genre_1 == "East Coast Hip Hop"
            )%>% group_by(Genre_1 ,Year)%>% summarise(
              Average.Score = mean(Average.Rating),
              count = n()
              )%>% ggplot(aes(x = Year,y = Average.Score, group = Genre_1, color = Genre_1)) + geom_line()+ theme(axis.text.x = element_text(angle = 90, size = 10))
```

Most genre's average scores over the years appear to be fairly consistent. Heavy Metal started out strong in 1970, but then suffered a steep decline by the end of the decade, plateauing around 3.75. Hard Rock had the lowest average score, and throughout the timeline barely broke 3.8 at its highest peak. Hard bop, Film Score, Singer/Songwriter and east coast hip hop seem to be oscillating around an average rating of 3.8.

```{r}
df_final %>% group_by(Genre_1)%>% 
  filter(
    Genre_1 == "Progressive Rock"|
    Genre_1 == "Singer/Songwriter"|
    Genre_1 == "Hard Bop"|
    Genre_1 == "Film Score"|
    Genre_1 == "Hard Rock"|
    Genre_1 == "Heavy Metal"|
    Genre_1 == "East Coast Hip Hop"
            )%>% summarise(
  Average.Score = mean(Average.Rating),
  count = n()
  ) %>% ggplot(aes(x=reorder(Genre_1,Average.Score),y=Average.Score, fill = Genre_1))+  
  geom_bar(stat="identity")+ 
  coord_flip() + 
  geom_text(aes(label=round(Average.Score, digit = 2), hjust = 1))+
  xlab("Genre")
```

The average scores for each genre are very similar which makes sense as we are looking at the top 5000 albums so they all should have fairly high scores. Film Score had the highest average score with 3.81, and Hard Rock had the lowest average score at 3.75.

```{r}
df_final%>% filter(Genre_1 == "Progressive Rock"|
            Genre_1 == "Singer/Songwriter"|
            Genre_1 == "Hard Bop"|
            Genre_1 == "Film Score"|
            Genre_1 == "Hard Rock"|
            Genre_1 == "Heavy Metal"|
            Genre_1 == "East Coast Hip Hop"
            )%>% group_by(Genre_1 ,Year)%>% summarise(
  count = n()
)%>% ggplot(aes(x = Year,y = count, group = Genre_1, color = Genre_1)) + geom_line()+ theme(axis.text.x = element_text(angle = 90, size = 10))
```

Looking at the count for the genres over time, it can be seen that from the beginning of the 50s to the end of the 60s Hard Bop Jazz was the dominating genre for music. The 70s is the Rock era with Progressive and Hard Rock occurring the most with the Singer/Songwriter genre close to the Hard Rock genre. The 80s had Heavy Metal as the genre that occurred the most. Finally, East Coast Hip Hop is the most occurring genre for the 90s. The Singer/Songwriter and Progressive rock genre came back a little in the early 2000s.

To answer the question: How do Genre's scores change over time? From the plot, the most popular genre's scores appear to oscillate over the years, with a fairly consistent average around 3.8 for all genres.


## Results 4 {data-link="Results 3"}

-What is the relationship between score and other variables, such as genre, descriptors, and artist? **Jasper**

By answering this question we will be able to have a better idea of how albums are scored and what the stronger criterion are that affect scores for each album. First we will look for the unique values of genres, descriptors and artists to identify how many of each category we have. Once that is accomplished, we will be using the dplyr library and grouping the data by genre to graph it with the ggplot library. Using these graphs we will be able to see how different categories affect score.

# First: Identify Question

# Question: What is the relationship between score and other variables, such as genre, descriptors, and artist?

### Find the Top 10 Most Common Genres, Descriptors, and Artists

```{r}
## Genre
df_final %>% group_by(Genre_1)%>% summarise(
  count = n()
)%>% arrange(desc(count))
```

-   Progressive Rock, Singer/Songwriter, Hard Bop, Film Score, Hard Rock, Heavy Metal, East Coast Hip Hop, Video Game Music, Alternative Rock, and Progressive Metal are the top 10 most common Genres.

```{r}
## Descriptor
df_final %>% group_by(Desc_1)%>% summarise(
  count = n()
)%>% arrange(desc(count))
```

-   Male Vocals, Instrumental, Energetic, Melodic, Melancholic, Female Vocals, Atmostpheric, Psychedelic, Passionate, and Dark are the top 10 most common Descriptors.

```{r}
## Artist
df_final %>% group_by(Artist.Name)%>% summarise(
  count = n()
)%>% arrange(desc(count))
```

-   Various Artists, Miles Davis, John Coltrane, Tom Waits, John Williams, Bob Dylan, The Beatles, Ennio Morricone, Herbie Hancock, Pharoah Sanders, and The Fall

### Average Score of the Top 10 Most Common Genres, Descriptors, and Artists

```{r}
## Mean of all albums in the list
mean(df_final$Average.Rating)
```

-   The average score for all albums in the list is 3.771304.

```{r}
## Average score of Top 10 Genres
df_final_genres <- df_final[df_final$Genre_1==c("Progressive Rock", "Singer/Songwriter", "Hard Bop", "Film Score", "Hard Rock", 
                                                "Heavy Metal", "East Coast Hip Hop", "Video Game Music", "Alternative Rock", "Progressive Metal"),]
mean(df_final_genres$Average.Rating)
```

-   The average score for albums with their primary genre in the top 10 most common genre is 3.775546.

```{r}
## Average score of Top 10 Descriptors
df_final_descriptors <- df_final[df_final$Desc_1==c("male vocals", "instrumental", "energetic", "melodic", "melancholic",
                                                    "female vocals", "atmospheric", "psychedelic", "passionate", "dark"),]
mean(df_final_descriptors$Average.Rating)
```

-   The average score for albums with their primary descriptor in the top 10 most common descriptors is 3.771124

```{r}
## Average score of Top 10 Artists
df_final_artists <- df_final[df_final$Artist.Name=="Various Artists" |  df_final$Artist.Name=="Miles Davis" | df_final$Artist.Name=="John Coltrane" |  df_final$Artist.Name=="Tom Waits" |
                            df_final$Artist.Name=="John Williams" | df_final$Artist.Name=="Bob Dylan" | df_final$Artist.Name=="The Beatles" |  df_final$Artist.Name=="Ennio Morricone" |
                            df_final$Artist.Name=="Herbie Hancock" | df_final$Artist.Name=="Pharoah Sanders" | df_final$Artist.Name=="The Fall",]
mean(df_final_artists$Average.Rating)
```

-   The average score for albums with their artists in the top 10 most common artists is 3.855435.

```{r}
## Average Score of all Top 10's
df_final_top10s <- df_final[df_final$Genre_1==c("Progressive Rock", "Singer/Songwriter", "Hard Bop", "Film Score", "Hard Rock", 
                                                "Heavy Metal", "East Coast Hip Hop", "Video Game Music", "Alternative Rock", "Progressive Metal") | 
                            df_final$Desc_1==c("male vocals", "instrumental", "energetic", "melodic", "melancholic",
                                                    "female vocals", "atmospheric", "psychedelic", "passionate", "dark") | 
                            df_final$Artist.Name=="Various Artists" |  df_final$Artist.Name=="Miles Davis" | df_final$Artist.Name=="John Coltrane" |  df_final$Artist.Name=="Tom Waits" |
                            df_final$Artist.Name=="John Williams" | df_final$Artist.Name=="Bob Dylan" | df_final$Artist.Name=="The Beatles" |  df_final$Artist.Name=="Ennio Morricone" |
                            df_final$Artist.Name=="Herbie Hancock" | df_final$Artist.Name=="Pharoah Sanders" | df_final$Artist.Name=="The Fall",]
mean(df_final_top10s$Average.Rating)
```

-   The average score for an album with any of the previous top 10 values included in the album is 3.7976607.

### Average Score of !Top 10 most common genres, descriptors, artists

```{r}
## Average score without top 10 most common Genres
df_final_genres2 <- df_final[!df_final$Genre_1==c("Progressive Rock", "Singer/Songwriter", "Hard Bop", "Film Score", "Hard Rock", 
                                                "Heavy Metal", "East Coast Hip Hop", "Video Game Music", "Alternative Rock", "Progressive Metal"),]
mean(df_final_genres2$Average.Rating)
```

-   The average score when the top 10 most common primary genres are removed is 3.771201.

```{r}
## Average score without top 10 most common Descriptors
df_final_descriptors2 <- df_final[!df_final$Desc_1==c("male vocals", "instrumental", "energetic", "melodic", "melancholic",
                                                    "female vocals", "atmospheric", "psychedelic", "passionate", "dark"),]
mean(df_final_descriptors2$Average.Rating)
```

-   The average score when the top 10 most common primary descriptors are removed is 3.771311

```{r}
## Average score without top 10 most common Artists
df_final_artists2 <- df_final[!df_final$Artist.Name=="Various Artists" &  !df_final$Artist.Name=="Miles Davis" & !df_final$Artist.Name=="John Coltrane" &  
                             !df_final$Artist.Name=="Tom Waits" & !df_final$Artist.Name=="John Williams" & !df_final$Artist.Name=="Bob Dylan" & 
                             !df_final$Artist.Name=="The Beatles" &  !df_final$Artist.Name=="Ennio Morricone" & !df_final$Artist.Name=="Herbie Hancock" & 
                             !df_final$Artist.Name=="Pharoah Sanders" & !df_final$Artist.Name=="The Fall",]
mean(df_final_artists2$Average.Rating)
```

-   The average score when the top 10 most common artists are removed is 3.768916.

```{r}
## Average score without all top 10's
df_final_top10s2 <- df_final[!df_final$Genre_1==c("Progressive Rock", "Singer/Songwriter", "Hard Bop", "Film Score", "Hard Rock", 
                                                "Heavy Metal", "East Coast Hip Hop", "Video Game Music", "Alternative Rock", "Progressive Metal") &
                             !df_final$Desc_1==c("male vocals", "instrumental", "energetic", "melodic", "melancholic",
                                                    "female vocals", "atmospheric", "psychedelic", "passionate", "dark") &
                             !df_final$Artist.Name=="Various Artists" &  !df_final$Artist.Name=="Miles Davis" & !df_final$Artist.Name=="John Coltrane" &  
                             !df_final$Artist.Name=="Tom Waits" & !df_final$Artist.Name=="John Williams" & !df_final$Artist.Name=="Bob Dylan" & 
                             !df_final$Artist.Name=="The Beatles" &  !df_final$Artist.Name=="Ennio Morricone" & !df_final$Artist.Name=="Herbie Hancock" & 
                             !df_final$Artist.Name=="Pharoah Sanders" & !df_final$Artist.Name=="The Fall",]
mean(df_final_top10s2$Average.Rating)
```

-   The average score when the top 10 most common artists, genres, and descriptors are removed is 3.768879.

### Correlation matrix to find what is most correlated with score

```{r}
dummydf <- df_final

dummydf$Genre_1 <- as.numeric(factor(dummydf$Genre_1))
dummydf$Desc_1 <- as.numeric(factor(dummydf$Desc_1))
dummydf$Artist.Name <-  as.numeric(factor(dummydf$Artist.Name))
dummydf$Year <-  as.numeric(factor(dummydf$Year))
dummydf$Number.of.Ratings <-  as.numeric(factor(dummydf$Number.of.Ratings))
dummydf$Number.of.Reviews <-  as.numeric(factor(dummydf$Number.of.Reviews))

dummydf2 <- subset(dummydf, select=c("Average.Rating","Artist.Name", "Genre_1", "Desc_1", "Year", "Number.of.Ratings", "Number.of.Reviews"))
results <- cor(dummydf2)
round(results,2)
```

-   Using our correlation matrix we can see that Number.of.Reviews is our strongest predictor of Average Rating with Number.of.Ratings close behind. If we were making a model to predict the average rating for an album we would include the variables: Number.of.Reviews, Number.of.Ratings, and Year.

# Conclusions {data-link="Conclusions"}

Give a quick summary of your work. Here is the place to be a bit critical and discuss potential limitations. Add a sentence on what else you would have liked to include in your data exploration if you had more time or more members in your team.

## Data source {.unnumbered}

Where does the data come from, who owns the data? Where are all the scripts that you need to clean the data?

## References {.unnumbered}

List all resources you used.
